Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-02 overleaf/related.tex
Committing changes to Related work and Results.

Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}.

While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and
AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits.
Our focus is at the intermediate level.

@@

Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
restrict flow between secure and insecure parts of the program.

@@

\paragraph{Classical HPC compilers.}
Automatic vectorization is a longstanding problem in high-performance computing (HPC).
There are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
MPC Source vectorization.
We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
result of the condition at runtime.
The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
can lead to \emph{slowdown}.
%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}

We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
results in HPC.
Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
Vectorization meshes in with those additional optimizations in non-trivial ways.

Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. 
%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

>>>>

Languages and compilers for secure computation have seen significant attention and advances in recent years. The early MPC compilers Fairplay~\cite{CCS:BenNisPin08},
and Sharemind~\cite{ESORICS:BogLauWil08} were followed by PICCO~\cite{CCS:ZhaSteBla13}, Obliv-C~\cite{Zahur:2015}, TinyGarble~\cite{SP:SHSSK15},
Wystiria~\cite{SP:RasHamHic14}, and others. A new generation of MPC compilers
includes SPDZ/SCALE-MAMBA/MP-SPDZ~\cite{Keller:2020} and the ABY/HyCC/MOTION~\cite{NDSS:DemSchZoh15,CCS:BDKKS18,Braun:2022} frameworks.
These two families are the state-of-the art and
are actively developed. Another recent development is Viaduct~\cite{Acay:2021}, a functional language and compiler that supports a range of secure computation
frameworks, including MPC and ZKP. Hastings et el. present a review of compiler frameworks~\cite{Hastings:2019}. We believe that our work is unique as 
as a facade between the high-level language and lower-level circuit protocols.

\begin{comment}
While each of these languages and compilers brings in new ideas and advances, none addresses the problem of ``circuit independent'' intermediate
representation and optimization. We envision a classical compiler structure: (1) a Wysteria, Viaduct, Obliv-C, or IMP Source front end, including rich type systems and
AST-level semantic analysis, compile into the MPC Source IR, (2) MPC Source-level optimizations take place, followed by (3) back-end compilers into circuits.
Our focus is at the intermediate level.

@@

Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
restrict flow between secure and insecure parts of the program.
\end{comment}

@@

\paragraph{Classical HPC compilers.}
Automatic vectorization is a longstanding problem in high-performance computing and
there are thousands of works in this area reflecting over 40 years of research. We presented a vectorization
algorithm for MPC Source, essentially extending classical loop vectorization \cite{Allen:1987}. In HPC vectorization, conditional control flow
presents a challenge --- one cannot estimate the cost of a schedule or vectorize branches in a straightforward manner --- in contrast to
MPC Source vectorization.
We view Karrenberg's work on Whole function vectorization~\cite{Karrenberg:2015} as most closely related to ours --- it linearizes the program and vectorizes
both branches of a conditional applying masking to avoid execution of the branch-not-taken code, and selection (similar to MUX) to select the correct value based on the
result of the condition at runtime.
The problem is that masking and selection, or more generally, handling control predicates~\cite{Benabderrahmane:2010,Karrenberg:2015},
can lead to \emph{slowdown}.
%As a result, vectorizing compilers and polyhedral compilers tend to focus on straight-line code due to the unpredictable
%overhead conditionals may introduce. \ana{add Louis noel pochet's benchmarks. citation on vectorizing compilers???}

We argue that vectorization over linear MPC Source is a different problem, one that warrants a new look, while drawing from
results in HPC.
%Since both branches of the conditional and the multiplexer \emph{always} execute, not only can we apply aggressive vectorization on linear code, but (perhaps more importantly)
%we can also build analytical models that accurately predict execution time. These models in turn would drive optimizations such as vectorization, protocol mixing, and others.
%Vectorization meshes in with those additional optimizations in non-trivial ways.
%Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
%and present novel challenges and opportunities for contribution. 
Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. 
%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

### 2022-05-02 overleaf/results.tex
Committing changes to Related work and Results.

   \begin{table*}[htbp]
       \centering
       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
       \label{table:metrics}
       \rowcolors{5}{}{gray!10}
       \tabcolsep3pt
       \begin{tabular}{lrrrrrrrrrrrr}
           \toprule
           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
          \midrule
            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
            \midrule
            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
            \midrule
            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
            \midrule
            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
            \midrule
            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
            \midrule
            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
            \midrule
            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
            \midrule
            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
            \midrule
            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
            \midrule
            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
            \midrule
            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
            \midrule
            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
            \midrule
            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
            \midrule
            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
            \midrule
            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
          \bottomrule
       \end{tabular}
       \label{table:metrics}
   \end{table*}

\begin{figure*}[htbp]
\centering
%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
\label{fig:graph_all_eval_time}
\end{figure*}

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection. These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of inner product and biometric matching. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, longest102 and max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. %Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time. 
Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments. As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively. %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CommunicationMiB}}
\caption{Communication Size of Benchmarks}
\label{fig:graph_comm_size}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-SetupTimesec}}
\caption{Setup Time of Benchmarks}
\label{fig:graph_setup_time}
\end{figure*}

@@

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop generated and ran {\sf result[i] = result[i-1] + mult\_unsimdified[i];} recall that the scalar expansion
is an artifact of our vectorization. We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. The poor performance of our loop is not a limitation of our approach, but it could be a 
limitation of our current implementation. Recall that the next phase of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) { result\_pref = result; } result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
will detect the associative accumulation in the scalar expansion code, it is realistic to expect that it will in the above only slightly more complex code. 
It still might lead to confusion in the static analysis as analysis on the AST is difficult (unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

We conjecture that MPC Source, a straight-forward representation, will not only enable detection of general associative loops, but also allow 
for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.

>>>>

{\small
   \begin{table*}[htbp]
       \centering
       \caption{Vectorized vs Non-Vectorized Comparison, times in seconds (in LAN setting where applicable), Communication in MiB, Numbers in 1000s, values rounded to nearest integer, benchmark names ending in {\em V} are vectorized.}
       \label{table:metrics}
       \rowcolors{5}{}{gray!10}
       \tabcolsep3pt
       \begin{tabular}{lrrrrrrrrrrrr}
           \toprule
           {} & \multicolumn{6}{c}{\textbf{GMW}} & \multicolumn{6}{c}{\textbf{BMR}} \\
            \cmidrule(r){2-7} \cmidrule(r){8-13} \\
            {\bf Benchmark} & Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. &  Online & Setup & \# Gates & Circ Gen & \# Msgs & Comm. \\
          \midrule
            Biometric Matching & 146 & 16 & 1,784 & 119 & 1,413 & 140 & 89 & 263 & 1,595 & 139 & 2,716 & 312\\
            Biometric Matching (V) & 12 & 4 & 34 & 2 & 28 & 14 & 2 & 13 & 30 & 4 & 61 & 130\\
            \midrule
            Convex Hull & 48 & 6 & 551 & 40 & 516 & 51 & 28 & 72 & 494 & 39 & 695 & 80\\
            Convex Hull (V) & 0 & 1 & 2 & 0 & 1 & 4 & 0 & 2 & 1 & 1 & 2 & 32\\
            \midrule
            Count 102 & 79 & 6 & 418 & 35 & 525 & 52 & 15 & 62 & 269 & 33 & 785 & 92\\
            Count 102 (V) & 71 & 5 & 316 & 24 & 332 & 34 & 11 & 30 & 167 & 16 & 304 & 59\\
            \midrule
            Count 10s & 79 & 6 & 419 & 35 & 525 & 52 & 14 & 62 & 270 & 33 & 785 & 92\\
            Count 10s (V) & 71 & 4 & 316 & 24 & 332 & 34 & 11 & 29 & 167 & 16 & 304 & 59\\
            \midrule
            Cryptonets (Max Pooling) & 50 & 11 & 688 & 46 & 554 & 55 & 36 & 89 & 608 & 51 & 898 & 110\\
            Cryptonets (Max Pooling) (V) & 1 & 1 & 7 & 1 & 2 & 5 & 2 & 4 & 7 & 2 & 12 & 49\\
            \midrule
            Database Join & 70 & 8 & 433 & 48 & 790 & 80 & 19 & 229 & 458 & 119 & 3,518 & 427\\
            Database Join (V) & 54 & 6 & 320 & 35 & 575 & 61 & 16 & 112 & 320 & 57 & 1,457 & 285\\
            \midrule
            Database Variance & 166 & 18 & 2,009 & 135 & 1,639 & 163 & 95 & 269 & 1,708 & 145 & 2,795 & 320\\
            Database Variance (V) & 37 & 6 & 321 & 24 & 334 & 43 & 10 & 30 & 170 & 13 & 178 & 141\\
            \midrule
            Histogram & 94 & 10 & 862 & 68 & 979 & 97 & 27 & 94 & 491 & 51 & 1,132 & 135\\
            Histogram (V) & 33 & 5 & 166 & 16 & 164 & 23 & 7 & 17 & 92 & 13 & 154 & 68\\
            \midrule
            Inner Product & 127 & 15 & 1,675 & 108 & 1,308 & 130 & 83 & 250 & 1,526 & 134 & 2,623 & 301\\
            Inner Product (V) & 16 & 5 & 158 & 12 & 165 & 25 & 6 & 18 & 83 & 7 & 86 & 127\\
            \midrule
            k-means & 108 & 12 & 1,333 & 88 & 1,090 & 108 & 63 & 185 & 1,141 & 99 & 1,958 & 225\\
            k-means (V) & 6 & 3 & 47 & 4 & 43 & 12 & 2 & 11 & 32 & 4 & 54 & 95\\
            \midrule
            Longest 102 & 93 & 7 & 650 & 52 & 713 & 71 & 26 & 93 & 475 & 49 & 1,091 & 128\\
            Longest 102 (V) & 169 & 6 & 544 & 41 & 519 & 53 & 25 & 60 & 369 & 33 & 605 & 95\\
            \midrule
            Max. Dist. b/w Symbols & 71 & 8 & 572 & 43 & 576 & 57 & 24 & 69 & 397 & 38 & 748 & 89\\
            Max. Dist. b/w Symbols (V) & 166 & 7 & 538 & 39 & 512 & 51 & 24 & 57 & 363 & 32 & 589 & 78\\
            \midrule
            Minimal Points & 35 & 5 & 458 & 31 & 369 & 37 & 24 & 46 & 401 & 26 & 347 & 40\\
            Minimal Points (V) & 0 & 1 & 1 & 0 & 1 & 3 & 0 & 1 & 1 & 0 & 1 & 16\\
            \midrule
            MNIST ReLU & 132 & 31 & 1,843 & 126 & 1,483 & 152 & 98 & 247 & 1,630 & 135 & 2,401 & 298\\
            MNIST ReLU (V) & 3 & 3 & 25 & 3 & 9 & 17 & 5 & 11 & 25 & 5 & 33 & 136\\
            \midrule
            Private Set Intersection & 95 & 9 & 558 & 59 & 1,049 & 104 & 22 & 186 & 591 & 96 & 2,639 & 302\\
            Private Set Intersection (V) & 1 & 2 & 1 & 2 & 1 & 8 & 1 & 8 & 2 & 4 & 2 & 122\\
          \bottomrule
       \end{tabular}
       \label{table:metrics}
   \end{table*}
}
\begin{figure*}[htbp]
\centering
%\resizebox{7in}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-OnlineSetupTimesec}}
\caption{Circuit Evaluation Time (Setup + Online) of Benchmarks}
\label{fig:graph_all_eval_time}
\end{figure*}

A detailed summary of the effects of vectorization on various benchmarks is presented in \cref{table:metrics}. We show circuit evaluation times in \cref{fig:graph_all_eval_time}. In terms of amenability to vectorization, we divide benchmarks into 3 categories: 1) {\it High:} these include convex hull, cryptonets max pooling, minimal points and private set intersection. These benchmarks are highly parallelizable and see 25x to 70x speedup in BMR, and 30x to 55x in GMW protocol. 2) {\it Medium:} these include biometric matching, DB Variance, histogram, inner product, k-means iteration and MNIST ReLU. These benchmarks have non-parallelizable phases e.g. the summing phase of inner product and biometric matching. Still, most computation is parallelizable and it results in speedup from 5x to 25x in BMR, and 2x to 25x in GMW protocol. 3) {\it Low:} these include the Database Join and the regular expression benchmarks (count 102, count 10, longest 102 and max distance between symbols). There is less parallelizable computation in these programs, thus the speedup is lower. We see a speedup from 1.1x to 2x in BMR. In GMW, DB Join, Count 102 and Count 10s see speedup from 1.1x to 1.3x. However, longest102 and max distance between symbols suffer a slowdown of 0.5x. There is opportunity for vectorization in these benchmarks according to our analytical model, particularly, a there is a large EQ that is vectorized, although a large portion of the loop cannot be vectorized. We observed that transformation to vectorized code increased multiplicative depth and, the negative effect of increased depth is more noticeable in a round-based protocol like GMW. The cause of the increase is not clear --- we conjecture that MOTION performs optimizations over the non-vectorized loop body that decreases depth; also, EQ is relatively inexpensive in Boolean GMW and BMR compared to ADD and MUL, which also de-emphasizes the benefit of vectorization. We propose a simple heuristic although we do leave all the benchmarks in the table: if the transformation increases circuit depth beyond some threshold (e.g. more than 10\% of the original circuit), we can reject the transformation. %Nevertheless, we show these graphs here for the sake of completeness and to highlight that vectorization is not always reduce run time. 
Note that in some settings it may still be desirable to vectorize e.g. in data constrained environments as communication is reduced for all benchmarks. (The report presents plots of communication, circuit generation time, number of gates, and separate online and setup time~\cite{Anon_TR}.)%As shown in \cref{fig:graph_comm_size}, vectorization results in reduced communication (fewer bits are transferred).

%We present evaluation for communication size in \cref{fig:graph_comm_size}, circuit generation time in \cref{fig:graph_circ_gen_time}, number of gates in \cref{fig:graph_total_gates} and online time and setup time in \cref{fig:graph_online_time} and \cref{fig:graph_setup_time} respectively. %\ishaq{Ana: Let me know if I should add more substance.} \ana{It's good!}

@@

\begin{comment}
\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-CommunicationMiB}}
\caption{Communication Size of Benchmarks}
\label{fig:graph_comm_size}
\end{figure*}

@@

\begin{figure*}[htbp]
\centering
\resizebox{0.7\textwidth}{!}{\input{graphs/lan/all-hist-SetupTimesec}}
\caption{Setup Time of Benchmarks}
\label{fig:graph_setup_time}
\end{figure*}
\end{comment}

@@

%We conclude the . 
Finally, we compared our automatically generated routine for Inner Product with the manually SIMD-ified MOTION-native routine in the distribution. 
We were surprised that we were an order of magnitude slower in Boolean GMW as our circuit ran a significantly larger 
number of communication rounds. Upon investigation, it turned out that the vectorized multiplications were essentially the same, 
however, our addition loop incurred significant cost (ADD is non-local and expensive in Boolean GMW). The MOTION-native loop ran 
{\sf result += mult\_unsimdified[i];} while our loop ran auto generated {\sf result[i] = result\_prev[i] + mult\_unsimdified[i];} (recall that scalar expansion
is an artifact of our vectorization). We rewrote the accumulation (manually, for testing purposes) and that led to the comparable speedup!

MOTION's compiler performs analysis that informs circuit generation and the example illustrates the power of the analysis. In the above example, 
MOTION does the standard divide-and-conquer accumulation. We conjectured that poor performance of our loop was due to a
limitation of our current implementation. Recall that the Phase 3 of Vectorization, in~\secref{sec:vectorization}, which we have not implemented yet, gets rid of redundant 
dimensions and generates {\sf if (i!=0) \{ result\_prev = result; \} result = result\_prev + mult\_unsimdified[i]}. And while it is unrealistic to expect that MOTION's static analysis 
will detect the associative accumulation in the scalar expansion code, one might expect that it will in the above slightly more complex code. 
Our experience indicates the limitation of AST-based analysis; because AST analysis is difficult it does not appear to detect the associative accumulation.
%(unfortunately, we ran out of time and did not experiment further).
%Interestingly, the non-vectorized version, which is the literal translation of inner product does not appear to benefit from any optimizations.

We conjecture that MPC Source, a straight-forward representation suitable for static analysis, will not only enable detection of general associative loops, but also allow 
for program synthesis to increase opportunities for divide-and-conquer parallelization~\cite{Farzan:2021}; we leave this as future work.
